---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    
    
---
Load Packages
```{r}
library(stargazer)
library(corrplot)
library(gplots)
library(dplyr)
library(ggplot2)
library(car)
library(plm)
library(tidymodels)
library(readr)
library(stringr)
library(tidyr)
library(sqldf)
library(data.table)
library(caret)
library(tidyverse)
library(plyr)
library(scales)
library(glmnet)
library(gsubfn)
library(proto)
library(rgl)
library(outliers)
library(doParallel)
library(ranger)
library(vip)
library(BBmisc)
library(randomForest)
library(rpart)
library(rpart.plot)
```

First the data downloaded from National animal nutrition program database is Loaded.
```{r}
# Load data ---------------------------------------------------------------
Performance      <- read.csv("~/BAM/Thesis/Test USDA/Performance.csv")
Performance      <- as.data.table(Performance)
dietarynutrients <- read.csv("~/BAM/Thesis/Test USDA/dietarynutrients.csv")
dietarynutrients <- as.data.table(dietarynutrients)
subjects <- read.csv("~/BAM/Thesis/Test USDA/subjects.csv")
subjects <- as.data.table(subjects)
```


###############################################################################
Data Preparation
##############################################################################

Information regarding the subject data is filtered and a combined data table is created in which all the required variables are collected.
```{r}
# include subject information into one combined data table----------------------

# create subject ID
subjects$id <- paste(subjects$PubID,subjects$TrialID)
subjects$id <- paste(subjects$id,subjects$TrtID)
subjects$id <- paste(subjects$id,subjects$SubjectID)

# Filter on Breed
Subjects_breed <- subjects %>%
  filter(VarName == "Subj_Breed")
names(Subjects_breed)[7] <- "Breed"
# Create a subset with only id and Breed
Subjects_breed    <- subset(Subjects_breed, select = c(7,12))

# Filter on sex
Subjects_Sex <- subjects %>%
  filter(VarName == "Subj_Sex")
names(Subjects_Sex)[7] <- "Sex"
# Create a subset with only id and Sex
Subjects_Sex    <- subset(Subjects_Sex, select = c(7,12))
# merge the subset with the combined data table
dt.combined <- merge(Subjects_breed, Subjects_Sex, by="id")

# Filter on LactDays
Subjects_LactDays <- subjects %>%
  filter(VarName == "Subj_LactDays")
names(Subjects_LactDays)[7] <- "LactDays"
# Create a subset with only id and LactDays
Subjects_LactDays    <- subset(Subjects_LactDays, select = c(7,12))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, Subjects_LactDays, by="id")

# Filter on PregDays
Subjects_PregDays <- subjects %>%
  filter(VarName == "Subj_PregDays")
names(Subjects_PregDays)[7] <- "PregDays"
# Create a subset with only id and PregDays
Subjects_PregDays    <- subset(Subjects_PregDays, select = c(7,12))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, Subjects_PregDays, by="id")

# make sure all numeric variables are set to numeric.
dt.combined$LactDays <- as.numeric(dt.combined$LactDays)
dt.combined$PregDays <- as.numeric(dt.combined$PregDays)
```

Information regarding the performance data is filtered and added to the combined data table in which all the required variables are collected.
```{r}
#Include subject performance data-----------------------------------------------

# Filter on Milk Production
Performance_Milk_prod    <- Performance[VarName == "Milk_Prod", ]
names(Performance_Milk_prod)[12] <- "Milk_prod_quant_KG_d"
# Create subject ID
Performance_Milk_prod$id <- paste(Performance_Milk_prod$PubID,Performance_Milk_prod$TrialID)
Performance_Milk_prod$id <- paste(Performance_Milk_prod$id,Performance_Milk_prod$TrtID)
Performance_Milk_prod$id <- paste(Performance_Milk_prod$id,Performance_Milk_prod$SubjectID)
# Create a subset with only id and Milk_prod_quant_KG_d
Performance_Milk_prod    <- subset(Performance_Milk_prod, select = c(1,12))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, Performance_Milk_prod, by="id")

# Filter on milk energy value
Performance_energy    <- Performance[VarName == "Milk_NE", ]
names(Performance_energy)[12] <- "Milk_Energy_Mcal_d"
# Create subject ID
Performance_energy$id <- paste(Performance_energy$PubID,Performance_energy$TrialID)
Performance_energy$id <- paste(Performance_energy$id,Performance_energy$TrtID)
Performance_energy$id <- paste(Performance_energy$id,Performance_energy$SubjectID)
# Create a subset with only id and Milk_Energy_Mcal_d
Performance_energy    <- subset(Performance_energy, select = c(1,12))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, Performance_energy, by="id")
```

Information regarding the performance data is filtered and added to the combined data table in which all the required variables are collected.
```{r}
# Include subject Dietarynutrients----------------------------------------------

# Create subject ID
dietarynutrients$id <- paste(dietarynutrients$PubID,dietarynutrients$TrialID)
dietarynutrients$id <- paste(dietarynutrients$id,dietarynutrients$TrtID)
dietarynutrients$id <- paste(dietarynutrients$id,dietarynutrients$SubjectID)

# Create a subset that includes only the values of the dietary ingredients
dietarynutrients<- subset(dietarynutrients, select = c(6:8,12))

# Filter on acid detergent fiber
nutrient_acid_detergent_fiber <- dietarynutrients[VarName == "Dt_ADF", ] 
names(nutrient_acid_detergent_fiber)[2] <- "Acid_detergent_fiber_perc_DM"
# Create a subset with only id and Acid detergent fiber
nutrient_acid_detergent_fiber <- subset(nutrient_acid_detergent_fiber, select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_acid_detergent_fiber, by="id")

# Filter on dietary ash
nutrient_ash <- dietarynutrients[VarName == "Dt_Ash", ]
names(nutrient_ash)[2] <- "Ash_perc_DM"
# Create a subset with only id and Ash
nutrient_ash <- subset(nutrient_ash, select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_ash, by="id")

# Filter on crude protein
nutrient_crude_protein <- dietarynutrients[VarName == "Dt_CP", ]
names(nutrient_crude_protein)[2] <- "Crude_protein_perc_DM"
# Create a subset with only id and Crude protein
nutrient_crude_protein <- subset(nutrient_crude_protein, select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_crude_protein, by="id")

# Filter on cellulose
nutrient_Cellulose <- dietarynutrients[VarName == "Dt_Cel", ]
names(nutrient_Cellulose)[2] <- "Cellulose_perc_DM"
# Create a subset with only id and cellulose
nutrient_Cellulose <- subset(nutrient_Cellulose, select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_Cellulose, by="id")

# Filter on crude fiber
nutrient_Crude_fiber <- dietarynutrients[VarName == "Dt_CF", ]
names(nutrient_Crude_fiber)[2] <- "Crude_fiber_perc_DM"
# Create a subset with only id and crude fiber
nutrient_Crude_fiber <- subset(nutrient_Crude_fiber, select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_Crude_fiber, by="id")

# Filter on dry matter
nutrient_dry_matter <- dietarynutrients[VarName == "Dt_DM", ]
names(nutrient_dry_matter)[2] <- "Dry_matter_g_d"
# Create a subset with only id and dry matter consumption
nutrient_dry_matter <- subset(nutrient_dry_matter, select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_dry_matter, by="id")

# Filter on hemicellulose
nutrient_Hemicellulose <- dietarynutrients[VarName == "Dt_HC", ] 
names(nutrient_Hemicellulose)[2] <- "Hemicellulose_perc_DM"
# Create a subset with only id and hemicellulose
nutrient_Hemicellulose <- subset(nutrient_Hemicellulose, select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_Hemicellulose, by="id")

# Filter on lignin
nutrient_lignin <- dietarynutrients[VarName == "Dt_Lig", ] 
names(nutrient_lignin)[2] <- "lignin_perc_DM"
# Create a subset with only id and lingin
nutrient_lignin <- subset(nutrient_lignin, select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_lignin, by="id")


# Filter on Neutral detergent fiber
nutrient_neutral_detergent_fiber <- dietarynutrients[VarName == "Dt_NDF", ] 
names(nutrient_neutral_detergent_fiber)[2] <- "Neutral_detergent_fiber_perc_DM"
# Create a subset with only id and Neutral detergent fiber
nutrient_neutral_detergent_fiber <- subset(nutrient_neutral_detergent_fiber, select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_neutral_detergent_fiber, by="id")


# Filter on Crude fat
nutrient_fat <- dietarynutrients[VarName == "Dt_EE", ] 
names(nutrient_fat)[2] <- "Crude_Fat_perc_DM"
# Create a subset with only id and organic matter
nutrient_fat  <- subset(nutrient_fat , select = c(4,2))
# merge the subset with the combined data table
dt.combined <- merge(dt.combined, nutrient_fat, by="id")

```

Cleaning and feature selection.

All data entries with a negative value are removed from the data set. 
These are likely to be entry errors since a animal cannot eat a negative amount 
of a nutritional ingredient. Also the single steer that is included is removed 
from the data set since this observation is not comparable to the rest of the 
data entries. The data set contains a couple of cows that have been in lactation
a long time. Research suggests that a full lactation cycle ideally takes around
year including the dry period. In practice the cycle can take a somewhat longer.
Because of this outliers that are in lactation longer than a year are removed
from the data set. There are three different breeds represented in the data set
namely Holstein, Jersey and Angus x hereford. These are not distributed equally
with most of the cows being Holstein cows. Since different breeds could have 
different nutritional requirements only Holstein cows are considered within
this research.
```{r}
#Clean data

# remove negative nutritional values
dt.combined <- dt.combined[dt.combined$Ash_perc_DM >=0,]
dt.combined <- dt.combined[dt.combined$Crude_Fat_perc_DM >=0,]
dt.combined <- dt.combined[dt.combined$Hemicellulose_perc_DM >=0,]
dt.combined <- dt.combined[dt.combined$lignin_perc_DM >=0,]


# remove cows that have been in lactation longer than a year
dt.combined <- dt.combined%>%
  filter(LactDays < 366)

# remove steer
dt.combined <- dt.combined%>%
  filter(Sex == "Female")

# Select only Holstein cows
dt.combined <- dt.combined%>%
  filter(Breed == "Holstein")

#count individual unique
Subject_id <- subset(subjects, select = c(5,12))
dt.combined_individual_cows <- merge(Subject_id, dt.combined, by="id")
dt.combined_individual_cows$SubjectID <- as.numeric(dt.combined_individual_cows$SubjectID)
dt.combined_individual_cows <- unique(dt.combined_individual_cows$SubjectID)
```


###############################################################################
Data exploration
##############################################################################

Stargazer is used to get a good summary of the overall data set
```{r}
stargazer(dt.combined, type = "latex")
stargazer(dt.combined, type = "text")
```

Distribution of Milk production
```{r}
#histogram distribution milk production
hist(dt.combined$Milk_prod_quant_KG_d,
     xlab = "Milk production per cow in Kg per day",
     main = "Distribution of milk production",
     breaks = sqrt(nrow(dt.combined))
)
```

Distribution of nutritional component values
```{r}
#histogram distribution of acid detergent fiber percentage in DM
hist(dt.combined$Acid_detergent_fiber_perc_DM,
     xlab = "Percentage of Acid detergent fiber in dry matter",
     main = "Distribution of Acid detergent fiber level in Dairy feed",
     breaks = sqrt(nrow(dt.combined))
)

#histogram distribution of Ash percentage in DM
hist(dt.combined$Ash_perc_DM,
     xlab = "Percentage of Ash in dry matter",
     main = "Distribution of Ash level in Dairy feed",
     breaks = sqrt(nrow(dt.combined))
)

#histogram distribution of acid detergent fiber percentage in DM
hist(dt.combined$Crude_protein_perc_DM,
     xlab = "Percentage of Crude protein in dry matter",
     main = "Distribution of Crude protein level in Dairy feed",
     breaks = sqrt(nrow(dt.combined))
)

#histogram distribution of acid detergent fiber percentage in DM
hist(dt.combined$Crude_protein_perc_DM,
     xlab = "Percentage of Crude fiber in dry matter",
     main = "Distribution of Crude fiber level in Dairy feed",
     breaks = sqrt(nrow(dt.combined))
)

#histogram distribution of acid detergent fiber percentage in DM
hist(dt.combined$Cellulose_perc_DM,
     xlab = "Percentage of cellulose in dry matter",
     main = "Distribution of cellulose level in Dairy feed",
     breaks = sqrt(nrow(dt.combined))
)

#histogram distribution of acid detergent fiber percentage in DM
hist(dt.combined$Hemicellulose_perc_DM,
     xlab = "Percentage of hemicellulose in dry matter",
     main = "Distribution of hemicellulose level in Dairy feed",
     breaks = sqrt(nrow(dt.combined))
)

#histogram distribution of acid detergent fiber percentage in DM
hist(dt.combined$lignin_perc_DM,
     xlab = "Percentage of lignin in dry matter",
     main = "Distribution of lignin level in Dairy feed",
     breaks = sqrt(nrow(dt.combined))
)

#histogram distribution of acid detergent fiber percentage in DM
hist(dt.combined$Neutral_detergent_fiber_perc_DM,
     xlab = "Percentage of neutral detergent fiber in dry matter",
     main = "Distribution of neutral detergent fiber level in Dairy feed",
     breaks = sqrt(nrow(dt.combined))
)

#histogram distribution of acid detergent fiber percentage in DM
hist(dt.combined$Crude_Fat_perc_DM,
     xlab = "Percentage of crude fat in dry matter",
     main = "Distribution of crude fat level in Dairy feed",
     breaks = sqrt(nrow(dt.combined))
)

```

A normalized version of the data set is create in order to better compare the
effects of different nutritional values.
```{r}
dt.combined_normalized <- normalize(dt.combined)
```

Create data table with only percentage of dry matter ingredients
```{r}
dt.nutrient_perc_DM <- subset(dt.combined_normalized, select = c(Milk_prod_quant_KG_d, Acid_detergent_fiber_perc_DM,Ash_perc_DM,Crude_protein_perc_DM,Crude_fiber_perc_DM, Cellulose_perc_DM,Hemicellulose_perc_DM,lignin_perc_DM, Neutral_detergent_fiber_perc_DM,  Crude_Fat_perc_DM))
```

OLS regression of Model 1 which includes only variables that are gives as a 
percentage of the dry matter.
```{r}
#OLS Regression with all nutrient variables
OLS_nutrient_variables <- lm(Milk_prod_quant_KG_d ~ ., data = dt.nutrient_perc_DM)
stargazer(OLS_nutrient_variables, type = "latex")
stargazer(OLS_nutrient_variables, type = "text")

```

Correlation of significant variables within the feed
```{r}
Corr_crude_fiber <- ggplot(dt.combined, aes(x = Crude_fiber_perc_DM, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Crude fiber percentage in Dry Matter") +
  ylab("Milk production")
Corr_crude_fiber

Corr_neutral_detergent_fiber <- ggplot(dt.combined, aes(x = Neutral_detergent_fiber_perc_DM, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Neutral Detergent fiber percentage in Dry Matter") +
  ylab("Milk production")
Corr_neutral_detergent_fiber

Corr_Crude_fat <- ggplot(dt.combined, aes(x = Crude_Fat_perc_DM, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Crude fat percentage in dry matter") +
  ylab("Milk production")
Corr_Crude_fat

Corr_Crude_protein <- ggplot(dt.combined, aes(x = Crude_protein_perc_DM, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Crude protein percentage in dry matter") +
  ylab("Milk production")
Corr_Crude_protein

Corr_hemicellulose <- ggplot(dt.combined, aes(x = Hemicellulose_perc_DM, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Hemicellulose percentage in dry matter") +
  ylab("Milk production")
Corr_hemicellulose

Corr_acid_detergent_fiber <- ggplot(dt.combined, aes(x = Acid_detergent_fiber_perc_DM, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Acid Detergent fiber percentage in Dry Matter") +
  ylab("Milk production")
Corr_acid_detergent_fiber


```

OLS regression of Model 2 which includes all available variables.
```{r}
#OLS Regression with all variables
dt.combined_ols <- select(dt.combined,4:17)
OLS_all_variables <- lm(Milk_prod_quant_KG_d ~ . , data = dt.combined_ols)
stargazer(OLS_nutrient_variables, OLS_all_variables, type = "latex")

dt.combined_ols_normalized <- select(dt.combined_normalized,4:17)
OLS_all_variables_normalized <- lm(Milk_prod_quant_KG_d ~ . , data = dt.combined_ols_normalized)
stargazer(OLS_all_variables_normalized, type = "text")
```

Correlation between Milk production and significant Variables
```{r}
# Correlation between Milk production and days in pregnancy
Corr_PregDays <- ggplot(dt.combined, aes(x = PregDays, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Days is pregnancy") +
  ylab("Milk production")
Corr_PregDays

# Correlation between Milk production and days in Milk
Corr_LactDays <- ggplot(dt.combined, aes(x = LactDays, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Days in Lactation") +
  ylab("Milk production")
Corr_LactDays

# Correlation between Milk production and net energy level
Corr_Energy <- ggplot(dt.combined, aes(x = Milk_Energy_Mcal_d, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Energy level for milk production in Mcal per day") +
  ylab("Milk production")
Corr_Energy

# Correlation between Milk production and Dry matter intake
Corr_Dry_matter <- ggplot(dt.combined, aes(x = Dry_matter_g_d, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Dry matter intake") +
  ylab("Milk production")
Corr_Dry_matter

Corr_Ash <- ggplot(dt.combined, aes(x = Ash_perc_DM, y = Milk_prod_quant_KG_d)) +
  geom_point() + geom_smooth(se = FALSE) + theme_minimal() +
  xlab("Ash percentage in Dry Matter") +
  ylab("Milk production")
Corr_Ash
```

Random forest regression in order to determine the importance of the variables
included in model 1.
```{r}
#Splitting the data into a training and test set
set.seed(2020)
data_split <- initial_split(data = dt.nutrient_perc_DM, prop = 0.7, 
                          strata = Milk_prod_quant_KG_d)
data_train <- training(data_split)
data_test <- testing(data_split)

#10-fold cross validation
set.seed(2020)
cv_folds <- vfold_cv(data_train, v = 10)

#Setting up the recipe
#str(data_train)
rf_recipe <- recipe(Milk_prod_quant_KG_d ~ ., data = data_train) 
rf_recipe

#Specify the random forest model with 1500 trees
tune_spec <- rand_forest(mtry = tune(), trees = 1500) %>%
  set_mode("regression") %>%
  set_engine("ranger")

#Setting up the workflow
tune_wf <- workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(tune_spec)
tune_wf

#Determine the metrics used
class_metrics <- metric_set(mae, rmse, rsq_trad)
registerDoParallel()

#Setting up the grid
set.seed(2020)
tune_res <- tune_grid(
  tune_wf,
  resamples = cv_folds,
  grid = tibble(mtry = 1:12),
  metrics = class_metrics
)
tune_res

tune_res %>%
  collect_metrics()

#Plotting the results of MAE, RMSE and R squared
tune_res %>%
  collect_metrics() %>%
  ggplot(aes(x = mtry, y = mean)) +
  geom_point() + geom_line() +
  facet_wrap(~ .metric, scales = "free_y") + theme_bw()

#Selecting the best model based on the MAE
best_mae <- select_best(tune_res, "mae")
best_mae
final_rf <- finalize_workflow(tune_wf, best_mae)

#Setting up the final workflow
set.seed(2020)
final_res <- final_rf %>%
  last_fit(data_split, metrics = class_metrics)

#Collecting the metrics
rf_metrics <- final_res %>%
  collect_metrics()
rf_metrics

plot_final_rf <- finalize_model(
  tune_spec,
  best_mae
)
#Plotting the variable important scores
plot_final_rf <- plot_final_rf %>% 
  set_engine("ranger", importance = "permutation") %>% 
  fit(Milk_prod_quant_KG_d ~ ., data = juice(prep(rf_recipe))) %>% vip(geom = "point") + theme_bw()
plot_final_rf
plot_final_rf[[1]][[2]]
```

Random forest regression in order to determine the importance of the variables
included in model 2.
```{r}
#normalize dt.combined
dt.combined_All_variables <- subset(dt.combined[,4:17])
dt.combined_All_variables <- normalize(dt.combined_All_variables)

#Splitting the data into a training and test set
set.seed(2020)
data_split <- initial_split(data = dt.combined_All_variables, prop = 0.7, 
                          strata = Milk_prod_quant_KG_d)
data_train <- training(data_split)
data_test <- testing(data_split)

#10-fold cross validation
set.seed(2020)
cv_folds <- vfold_cv(data_train, v = 10)


#Setting up the recipe
#str(data_train)
rf_recipe <- recipe(Milk_prod_quant_KG_d ~ ., data = data_train) #%>% update_role(rownames, new_role = "ID")
rf_recipe

#Specify the random forest model with 1500 trees
tune_spec <- rand_forest(mtry = tune(), trees = 1500) %>%
  set_mode("regression") %>%
  set_engine("ranger")

#Setting up the workflow
tune_wf <- workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(tune_spec)
tune_wf

#Determine the metrics used
class_metrics <- metric_set(mae, rmse, rsq_trad)
registerDoParallel()

#Setting up the grid
set.seed(2020)
tune_res <- tune_grid(
  tune_wf,
  resamples = cv_folds,
  grid = tibble(mtry = 1:12),
  metrics = class_metrics
)
tune_res

tune_res %>%
  collect_metrics()

#Plotting the results of MAE, RMSE and R squared
tune_res %>%
  collect_metrics() %>%
  ggplot(aes(x = mtry, y = mean)) +
  geom_point() + geom_line() +
  facet_wrap(~ .metric, scales = "free_y") + theme_bw()



#Selecting the best model based on the MAE
best_mae <- select_best(tune_res, "mae")
best_mae
final_rf <- finalize_workflow(tune_wf, best_mae)

#Setting up the final workflow
set.seed(2020)
final_res <- final_rf %>%
  last_fit(data_split, metrics = class_metrics)

#Collecting the metrics
rf_metrics <- final_res %>%
  collect_metrics()
rf_metrics

plot_final_rf <- finalize_model(
  tune_spec,
  best_mae
)

#Plotting the variable important scores
plot_final_rf <- plot_final_rf %>% 
  set_engine("ranger", importance = "permutation") %>% 
  fit(Milk_prod_quant_KG_d ~ ., data = juice(prep(rf_recipe))) %>% vip(geom = "point") + theme_bw()
plot_final_rf
plot_final_rf[[1]][[2]]

```


Create subsets based on the phase of lactation. All overservations with a Lactsdays
value lower than 101 are classified as early lactation. Observations ith a 
Lactdays value ranging from 101 to 200 are classified as mid-lactation. Observations
with a LactDays value higher than 200 will be classified as late-lactation.
```{r}
# create subset of cows that are in early lactation
dt.combined_early_lactation <- dt.combined %>%
  filter(LactDays < 101)

# select only the nutritional values given as a percentage of the dry matter
dt.nutrient_early_lactation <- subset(dt.combined_early_lactation, select = c(Milk_prod_quant_KG_d, Acid_detergent_fiber_perc_DM,Ash_perc_DM,Crude_protein_perc_DM,Crude_fiber_perc_DM, Cellulose_perc_DM,Hemicellulose_perc_DM,lignin_perc_DM, Neutral_detergent_fiber_perc_DM, Crude_Fat_perc_DM))

# normalize data frame
dt.nutrient_early_lactation_normalized <- normalize(dt.nutrient_early_lactation)

# create subset of cows that are in early in mid-lactation
dt.combined_mid_lactation <- dt.combined %>%
  filter(LactDays > 100 & LactDays < 201)

dt.nutrient_mid_lactation <- subset(dt.combined_mid_lactation, select = c(Milk_prod_quant_KG_d, Acid_detergent_fiber_perc_DM,Ash_perc_DM,Crude_protein_perc_DM,Crude_fiber_perc_DM, Cellulose_perc_DM,Hemicellulose_perc_DM,lignin_perc_DM, Neutral_detergent_fiber_perc_DM,  Crude_Fat_perc_DM))

# normalize data frame
dt.nutrient_mid_lactation_normalized <- normalize(dt.nutrient_mid_lactation)

# create subset of cows that are in early in mid-lactation
dt.combined_late_lactation <- dt.combined %>%
  filter(LactDays > 201)

dt.nutrient_late_lactation <- subset(dt.combined_late_lactation, select = c(Milk_prod_quant_KG_d, Acid_detergent_fiber_perc_DM,Ash_perc_DM,Crude_protein_perc_DM,Crude_fiber_perc_DM, Cellulose_perc_DM,Hemicellulose_perc_DM,lignin_perc_DM, Neutral_detergent_fiber_perc_DM,  Crude_Fat_perc_DM))

# normalize data frame
dt.nutrient_late_lactation_normalized <- normalize(dt.nutrient_late_lactation)
```


Distribution of milk production in different lactation phases.
```{r}
#histogram distribution milk production for cows in early lactation
hist(dt.nutrient_early_lactation$Milk_prod_quant_KG_d,
     xlab = "Milk production per cow in Kg per day",
     main = "Distribution of milk production during the early lactation phase",
     breaks = sqrt(nrow(dt.combined))
)

#histogram distribution milk production for cows in mid-lactation
hist(dt.nutrient_mid_lactation$Milk_prod_quant_KG_d,
     xlab = "Milk production per cow in Kg per day",
     main = "Distribution of milk production during the mid-lactation phase",
     breaks = sqrt(nrow(dt.combined))
)


#histogram distribution milk production for cows in late lactation
hist(dt.nutrient_late_lactation$Milk_prod_quant_KG_d,
     xlab = "Milk production per cow in Kg per day",
     main = "Distribution of milk production during the late lactation phase",
     breaks = sqrt(nrow(dt.combined))
)
```


Random forest regression in order to determine the importance of the variables
included in model 1 for cows that are in early lactation.
```{r}
#Splitting the data into a training and test set
set.seed(2020)
data_split <- initial_split(data = dt.nutrient_early_lactation_normalized, prop = 0.7, 
                          strata = Milk_prod_quant_KG_d)
data_train <- training(data_split)
data_test <- testing(data_split)

#10-fold cross validation
set.seed(2020)
cv_folds <- vfold_cv(data_train, v = 10)


#Setting up the recipe
#str(data_train)
rf_recipe <- recipe(Milk_prod_quant_KG_d ~ ., data = data_train) #%>% update_role(rownames, new_role = "ID")
rf_recipe

#Specify the random forest model with 1500 trees
tune_spec <- rand_forest(mtry = tune(), trees = 1500) %>%
  set_mode("regression") %>%
  set_engine("ranger")

#Setting up the workflow
tune_wf <- workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(tune_spec)
tune_wf

#Determine the metrics used
class_metrics <- metric_set(mae, rmse, rsq_trad)
registerDoParallel()

#Setting up the grid
set.seed(2020)
tune_res <- tune_grid(
  tune_wf,
  resamples = cv_folds,
  grid = tibble(mtry = 1:12),
  metrics = class_metrics
)
tune_res

tune_res %>%
  collect_metrics()

#Plotting the results of MAE, RMSE and R squared
tune_res %>%
  collect_metrics() %>%
  ggplot(aes(x = mtry, y = mean)) +
  geom_point() + geom_line() +
  facet_wrap(~ .metric, scales = "free_y") + theme_bw()


#Selecting the best model based on the MAE
best_mae <- select_best(tune_res, "mae")
best_mae
final_rf <- finalize_workflow(tune_wf, best_mae)

#Setting up the final workflow
set.seed(2020)
final_res <- final_rf %>%
  last_fit(data_split, metrics = class_metrics)

#Collecting the metrics
rf_metrics <- final_res %>%
  collect_metrics()
rf_metrics

plot_final_rf <- finalize_model(
  tune_spec,
  best_mae
)

#Plotting the variable important scores
plot_final_rf <- plot_final_rf %>% 
  set_engine("ranger", importance = "permutation") %>% 
  fit(Milk_prod_quant_KG_d ~ ., data = juice(prep(rf_recipe))) %>% vip(geom = "point") + theme_bw()
plot_final_rf
plot_final_rf[[1]][[2]]
```

Random forest regression in order to determine the importance of the variables
included in model 1 for cows that are in mid lactation.
```{r}
#Splitting the data into a training and test set
set.seed(2020)
data_split <- initial_split(data = dt.nutrient_mid_lactation_normalized, prop = 0.7, 
                          strata = Milk_prod_quant_KG_d)
data_train <- training(data_split)
data_test <- testing(data_split)

#10-fold cross validation
set.seed(2020)
cv_folds <- vfold_cv(data_train, v = 10)


#Setting up the recipe
#str(data_train)
rf_recipe <- recipe(Milk_prod_quant_KG_d ~ ., data = data_train) #%>% update_role(rownames, new_role = "ID")
rf_recipe

#Specify the random forest model with 1500 trees
tune_spec <- rand_forest(mtry = tune(), trees = 1500) %>%
  set_mode("regression") %>%
  set_engine("ranger")

#Setting up the workflow
tune_wf <- workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(tune_spec)
tune_wf

#Determine the metrics used
class_metrics <- metric_set(mae, rmse, rsq_trad)
registerDoParallel()

#Setting up the grid
set.seed(2020)
tune_res <- tune_grid(
  tune_wf,
  resamples = cv_folds,
  grid = tibble(mtry = 1:12),
  metrics = class_metrics
)
tune_res

tune_res %>%
  collect_metrics()

#Plotting the results of MAE, RMSE and R squared
tune_res %>%
  collect_metrics() %>%
  ggplot(aes(x = mtry, y = mean)) +
  geom_point() + geom_line() +
  facet_wrap(~ .metric, scales = "free_y") + theme_bw()


#Selecting the best model based on the MAE
best_mae <- select_best(tune_res, "mae")
best_mae
final_rf <- finalize_workflow(tune_wf, best_mae)

#Setting up the final workflow
set.seed(2020)
final_res <- final_rf %>%
  last_fit(data_split, metrics = class_metrics)

#Collecting the metrics
rf_metrics <- final_res %>%
  collect_metrics()
rf_metrics

plot_final_rf <- finalize_model(
  tune_spec,
  best_mae
)

#Plotting the variable important scores
plot_final_rf <- plot_final_rf %>% 
  set_engine("ranger", importance = "permutation") %>% 
  fit(Milk_prod_quant_KG_d ~ ., data = juice(prep(rf_recipe))) %>% vip(geom = "point") + theme_bw()
plot_final_rf
plot_final_rf[[1]][[2]]
```

Random forest regression in order to determine the importance of the variables
included in model 1 for cows that are in late lactation.
```{r}
#Splitting the data into a training and test set
set.seed(2020)
data_split <- initial_split(data = dt.nutrient_late_lactation_normalized, prop = 0.7, 
                          strata = Milk_prod_quant_KG_d)
data_train <- training(data_split)
data_test <- testing(data_split)

#10-fold cross validation
set.seed(2020)
cv_folds <- vfold_cv(data_train, v = 10)


#Setting up the recipe
#str(data_train)
rf_recipe <- recipe(Milk_prod_quant_KG_d ~ ., data = data_train) #%>% update_role(rownames, new_role = "ID")
rf_recipe

#Specify the random forest model with 1500 trees
tune_spec <- rand_forest(mtry = tune(), trees = 1500) %>%
  set_mode("regression") %>%
  set_engine("ranger")

#Setting up the workflow
tune_wf <- workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(tune_spec)
tune_wf

#Determine the metrics used
class_metrics <- metric_set(mae, rmse, rsq_trad)
registerDoParallel()

#Setting up the grid
set.seed(2020)
tune_res <- tune_grid(
  tune_wf,
  resamples = cv_folds,
  grid = tibble(mtry = 1:12),
  metrics = class_metrics
)
tune_res

tune_res %>%
  collect_metrics()

#Plotting the results of MAE, RMSE and R squared
tune_res %>%
  collect_metrics() %>%
  ggplot(aes(x = mtry, y = mean)) +
  geom_point() + geom_line() +
  facet_wrap(~ .metric, scales = "free_y") + theme_bw()


#Selecting the best model based on the MAE
best_mae <- select_best(tune_res, "mae")
best_mae
final_rf <- finalize_workflow(tune_wf, best_mae) 

#Setting up the final workflow
set.seed(2020)
final_res <- final_rf %>%
  last_fit(data_split, metrics = class_metrics)

#Collecting the metrics
rf_metrics <- final_res %>%
  collect_metrics()
rf_metrics

plot_final_rf <- finalize_model(
  tune_spec,
  best_mae
)

#Plotting the variable important scores
plot_final_rf <- plot_final_rf %>% 
  set_engine("ranger", importance = "permutation") %>% 
  fit(Milk_prod_quant_KG_d ~ ., data = juice(prep(rf_recipe))) %>% vip(geom = "point") + theme_bw()
plot_final_rf
plot_final_rf[[1]][[2]]
```


